---
title: "Maryland Road Crash Analysis: Aggregation"
author: "Arpit Jayesh Vora"
date: "21 April 2018"
output: html_document
---

## Aggregation
- Aggregation was important in this case to cluster the closely related (similar attibute) and remove redundant attributes. This will be beneficial in pre - processing stage which is done in Python.
- Aggregation is done based on related attribute values (REF: Data Dictionary) and also if the value significance is low (very less #records)
- Assumption: All NA's were replaced with code '88' = 'Others'
- Aggregation of Weather Contrib Type

```{r,eval=FALSE}
#--------------------------all contrib type aggregation-------------------------------------
table(Complete2016_v1$Weather_CONTRIB_CODE)
Complete2016_v1$Weather_CONTRIB_CODE[Complete2016_v1$Weather_CONTRIB_CODE == 44] <- 43
Complete2016_v1$Weather_CONTRIB_CODE[Complete2016_v1$Weather_CONTRIB_CODE == 45] <- 42
Complete2016_v1$Weather_CONTRIB_CODE[Complete2016_v1$Weather_CONTRIB_CODE == 47] <- 41
Complete2016_v1$Weather_CONTRIB_CODE[Complete2016_v1$Weather_CONTRIB_CODE == 82.88] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$Weather_CONTRIB_CODE[is.na(Complete2016_v1$Weather_CONTRIB_CODE)]<-88
```
- Weather Code aggregation (numeric values are the code for certain values)
```{r,eval=FALSE}
#--------------------------weather code aggregation-----------------------------------------
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 10.04] <- 5
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 11.88] <- 5
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 12.04] <- 5
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 7.01] <- 3
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 8.04] <- 3
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 9.04] <- 3
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 99] <- 88
Complete2016_v1$WEATHER_CODE[Complete2016_v1$WEATHER_CODE == 6.01] <- 6
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$WEATHER_CODE[is.na(Complete2016_v1$WEATHER_CODE)]<-88
```
- Surf Condition Code aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Surf condition code--------------------------------
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 4] <- 3
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 99] <- 88
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 5] <- 88
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 6.03] <- 88
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 7.01] <- 88
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 8.05] <- 88
Complete2016_v1$SURF_COND_CODE[Complete2016_v1$SURF_COND_CODE == 9.88] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$SURF_COND_CODE[is.na(Complete2016_v1$SURF_COND_CODE)]<-88
```
- Road Condition Code aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation road condition code--------------------------------
for (i in 3:8) {
  Complete2016_v1$RD_COND_CODE[Complete2016_v1$RD_COND_CODE == i] <- 2
}
Complete2016_v1$RD_COND_CODE[Complete2016_v1$RD_COND_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$RD_COND_CODE[is.na(Complete2016_v1$RD_COND_CODE)]<-88
```
- Injury Severe Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation injury severe code--------------------------------
for (i in 3:5) {
  Complete2016_v1$INJ_SEVER_CODE[Complete2016_v1$INJ_SEVER_CODE == i] <- 2
}
Complete2016_v1$INJ_SEVER_CODE[Complete2016_v1$INJ_SEVER_CODE == 1] <- 0
Complete2016_v1$INJ_SEVER_CODE[Complete2016_v1$INJ_SEVER_CODE == 2] <- 1
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$INJ_SEVER_CODE[is.na(Complete2016_v1$INJ_SEVER_CODE)]<-88
```
- Equip Prob Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation equip prob code-----------------------------------
Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == 11] <- 2
Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == 13] <- 2
Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == 31] <- 2
Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == 42] <- 2
for (i in 44:47) {
  Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == i] <- 2
}
Complete2016_v1$EQUIP_PROB_CODE[Complete2016_v1$EQUIP_PROB_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$EQUIP_PROB_CODE[is.na(Complete2016_v1$EQUIP_PROB_CODE)]<-88
```
- Safety Equi Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation safety equip code-----------------------------------
Complete2016_v1$SAF_EQUIP_CODE[Complete2016_v1$SAF_EQUIP_CODE == 12] <- 11
Complete2016_v1$SAF_EQUIP_CODE[Complete2016_v1$SAF_EQUIP_CODE == 13] <- 11
Complete2016_v1$SAF_EQUIP_CODE[Complete2016_v1$SAF_EQUIP_CODE == 22] <- 21
Complete2016_v1$SAF_EQUIP_CODE[Complete2016_v1$SAF_EQUIP_CODE == 23] <- 21
Complete2016_v1$SAF_EQUIP_CODE[Complete2016_v1$SAF_EQUIP_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$SAF_EQUIP_CODE[is.na(Complete2016_v1$SAF_EQUIP_CODE)]<-88
```
- Eject Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Eject code-----------------------------------
Complete2016_v1$EJECT_CODE[Complete2016_v1$EJECT_CODE == 3] <- 2
Complete2016_v1$EJECT_CODE[Complete2016_v1$EJECT_CODE == 4] <- 2
Complete2016_v1$EJECT_CODE[Complete2016_v1$EJECT_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$EJECT_CODE[is.na(Complete2016_v1$EJECT_CODE)]<-88
```
- Person Condition Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Person Condition code-----------------------------------
for (i in 3:8) {
  Complete2016_v1$CONDITION_CODE[Complete2016_v1$CONDITION_CODE == i] <- 2
}
Complete2016_v1$CONDITION_CODE[Complete2016_v1$CONDITION_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$CONDITION_CODE[is.na(Complete2016_v1$CONDITION_CODE)]<-88
```
- Person Contrib Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Person contrib code-----------------------------------
for (i in 2:4) {
  Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == i] <- 1
}
for (i in 8:17) {
  Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == i] <- 8
}
for (i in 20:28) {
  Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == i] <- 20
}
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 70.88] <- 20
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 73.88] <- 20
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 74.88] <- 20
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 75.88] <- 20

Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 29] <- 88
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 38.88] <- 88
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 39.88] <- 88
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 40.88] <- 88
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 60.88] <- 88
Complete2016_v1$Person_CONTRIB_CODE[Complete2016_v1$Person_CONTRIB_CODE == 68.88] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$Person_CONTRIB_CODE[is.na(Complete2016_v1$Person_CONTRIB_CODE)]<-88
```
- Damage Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Damage code-----------------------------------
Complete2016_v1$DAMAGE_CODE[Complete2016_v1$DAMAGE_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$DAMAGE_CODE[is.na(Complete2016_v1$DAMAGE_CODE)]<-88
```
- Light Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Light code-----------------------------------
Complete2016_v1$LIGHT_CODE[Complete2016_v1$LIGHT_CODE == 5.02] <- 5
Complete2016_v1$LIGHT_CODE[Complete2016_v1$LIGHT_CODE == 6.02] <- 6
Complete2016_v1$LIGHT_CODE[Complete2016_v1$LIGHT_CODE == 7.88] <- 7
Complete2016_v1$LIGHT_CODE[Complete2016_v1$LIGHT_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$LIGHT_CODE[is.na(Complete2016_v1$LIGHT_CODE)]<-88
```
- Junction Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Junction code-----------------------------------
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 5.01] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 3] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 6.01] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 7.01] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 8.04] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 9.04] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 10.04] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 11.04] <- 2
Complete2016_v1$JUNCTION_CODE[Complete2016_v1$JUNCTION_CODE == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$JUNCTION_CODE[is.na(Complete2016_v1$JUNCTION_CODE)]<-88
```
- Harm Event Code #1 aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Harm Event code #1-----------------------------------
for (i in 3:7) {
  Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == i] <- 3
}
for (i in 11:15) {
  Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == i] <- 11
}
for (i in 17:20) {
  Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == i] <- 17
}
Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == 17] <- 88
Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == 21.15] <- 88
Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == 22.15] <- 88
Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == 23.15] <- 88
Complete2016_v1$HARM_EVENT_CODE1[Complete2016_v1$HARM_EVENT_CODE1 == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$HARM_EVENT_CODE1[is.na(Complete2016_v1$HARM_EVENT_CODE1)]<-88
```
- Harm Event Code #2 aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Harm Event code #2-----------------------------------
for (i in 3:7) {
  Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == i] <- 3
}
for (i in 11:15) {
  Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == i] <- 11
}
for (i in 17:20) {
  Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == i] <- 17
}
Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == 17] <- 88
Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == 21.15] <- 88
Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == 22.15] <- 88
Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == 23.15] <- 88
Complete2016_v1$HARM_EVENT_CODE2[Complete2016_v1$HARM_EVENT_CODE2 == 99] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$HARM_EVENT_CODE2[is.na(Complete2016_v1$HARM_EVENT_CODE2)]<-88
```
- Body Type Code Aggregation
```{r,eval=FALSE}
#-----------------------------Aggregation Body code-----------------------------------------
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 99] <- 88

for (i in 13:18) {
  Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == i] <- 13
}
for (i in 3:12) {
  Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == i] <- 3
}
for (i in 19:22) {
  Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == i] <- 3
}
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 22.05] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 23.08] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 24.88] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 25.88] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 26.88] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 27.88] <- 3
Complete2016_v1$BODY_TYPE_CODE[Complete2016_v1$BODY_TYPE_CODE == 28] <- 3
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$BODY_TYPE_CODE[is.na(Complete2016_v1$BODY_TYPE_CODE)]<-88
```
- Area Damage Code
```{r,eval=FALSE}
#-----------------------------Aggregation Area Damage code-----------------------------------
Complete2016_v1$AREA_DAMAGED_CODE_IMP1[Complete2016_v1$AREA_DAMAGED_CODE_IMP1 == 99] <- 88
Complete2016_v1$AREA_DAMAGED_CODE_IMP1[Complete2016_v1$AREA_DAMAGED_CODE_IMP1 == 98] <- 88
#--------------------------Replacing NA's---------------------------------------------------
Complete2016_v1$AREA_DAMAGED_CODE_IMP1[is.na(Complete2016_v1$AREA_DAMAGED_CODE_IMP1)]<-88
```
